MaBoSS for HPC environments: Implementations of the continuous time Boolean model simulator for large CPU clusters and GPU accelerators

### 2024-02-11 Article.tex
add some refs

There are multiple ways to tackle the problem of bad cache performance of the expression tree. The first one is to use a different data structure that is more cache-friendly, such as the van Emde Boas tree layout. In our approach, we take one step further by utilizing the \emph{runtime compilation of GPU code}. Using this technique, instead of building the expression tree for the Boolean formulas, we compile the formulas into a native binary code that can be executed directly on the GPU. The main advantage of this approach is that the formulas are encoded in the instructions instead of being stored in the memory. This way, the formulas would not need to be fetched from memory multiple times during the application runtime. The other advantage is that the compiler can optimize the GPU code to utilize the GPU hardware more efficiently.

@@

For the benchmarks, we used $3$ real-world models with varying model sizes spanning from $10$ nodes to $133$. In order to test the scalability of the GPU and MPI implementation, we also created a synthetic models with up to $1000$ nodes. Synthetic models are designed in a way such that the length of each trajectory is predictable, and the models have no stable states. Also, the number of non-internal nodes is kept low (10 nodes) to enable the usage of the histogram optimization.

>>>>

There are multiple ways to tackle the problem of bad cache performance of the expression tree. The first one is to use a different data structure that is more cache-friendly, such as the van Emde Boas tree layout~\cite{van1975preserving}. In our approach, we take one step further by utilizing the \emph{runtime compilation of GPU code}. Using this technique, instead of building the expression tree for the Boolean formulas, we compile the formulas into a native binary code that can be executed directly on the GPU. The main advantage of this approach is that the formulas are encoded in the instructions instead of being stored in the memory. This way, the formulas would not need to be fetched from memory multiple times during the application runtime. The other advantage is that the compiler can optimize the GPU code to utilize the GPU hardware more efficiently.

@@

For the benchmarks, we used $3$ real-world models with varying model sizes spanning from $10$ nodes to $133$ (\texttt{cellcycle}~\cite{faure2006cellcycle}, \texttt{sizek}~\cite{sizek2019boolean} and \texttt{Montagud}~\cite{montagud2022prostate}). In order to test the scalability of the GPU and MPI implementation, we also created a synthetic models with up to $1000$ nodes. Synthetic models are designed in a way such that the length of each trajectory is predictable, and the models have no stable states. Also, the number of non-internal nodes is kept low (10 nodes) to enable the usage of the histogram optimization.
